### 2022-03-15 omicron-generation.tex
some more clarification

Serial interval data are taken from the supplementary material of \cite{backer2021omicron}.
The data are aggregated by the length of the serial interval in days and do not include additional individual-level information, such as exposure dates or symptom onset dates.
The data consists of 2529 transmission pairs and are further stratified by the presence of S gene target failure, week of infectors' symptom onset date (week 50 and 51), and the type of transmission (within or between households).
For simplicity, we refer to transmission pairs with and without S gene target failure as Omicron and Delta transmission pairs, respectively.
See original article for details of data collection.

@@

In practice, incubation periods (and other epidemiological delays) can be measured in two ways.
The forward incubation periods are measured from a cohort of individuals who were infected at the same time. 
We expect this forward incubation-period distribution $f_I(\tau)$ to remain constant over the course of an epidemic and provide reliable estimates of the distribution across individuals. 
Backward incubation periods are measured from a cohort of individuals who developed symptoms at the same time. 
We expect the backward incubation-period distribution to be sensitive to epidemic dynamics.
The difference arises because forward incubation periods look forward from the reference point towards symptom development, which is an individual-level process, while backward incubation periods look backwards towards an infection event, which requires an interaction with an infectious individual.

@@

We adjust for observed growth rate of each variant by taking the backward incubation-period distributions estimated by \cite{backer2021omicron} and calculating the corresponding forward incubation-period distributions by inverting \eref{backward}:
\begin{equation}
f_I(\tau) = \frac{\exp(r\tau) b_I(\tau) }{\int_0^\infty \exp(rx) b_I(x)\dx}.
\label{eq:forward}
\end{equation}
We model the backward incubation-period distribution $b_I(\tau)$ using a Weibull distribution based on the assumptions of \cite{backer2021omicron}.
To account for uncertainties in the original parameter estimates, we rely on a sampling scheme, similar to the one we used for the growth rate analysis (in Section 2.2).
First, we approximate the previously inferred posterior distributions of the shape and scale parameters of the Weibull distribution using a lognormal distribution---we parameterize the lognormal distribution such that (i) its median matches the median of the posterior distributions and (ii) the probability that a random variable following the specified Gamma distribution falls between the lower and upper credible limits matches 95\% \citep{park2020reconciling}.
We draw 1000 samples of the shape and scale parameters (for the backward distribution $b_I(\tau)$) from the specified lognormal distributions and estimate the corresponding forward distribution using \eref{forward}.
We take 95\% equi-tailed quantiles to obtain 95\% confidence intervals.

>>>>

Serial interval data are taken from the supplementary material of \cite{backer2021omicron}.
The data are aggregated by the length of the serial interval in days and do not include additional individual-level information, such as exposure dates or symptom onset dates.
The data consists of 2529 transmission pairs and are further stratified by the presence of S gene target failure (SGTF), week of infectors' symptom onset date (weeks 50 and 51), and the type of transmission (within or between households).
In the main text, we combine data from weeks 50 and 51 and instead present a stratified analysis in Supplementary Material.
For simplicity, we refer to transmission pairs with and without S gene target failure as Omicron and Delta transmission pairs, respectively.
Incubation period data are not publicly available with the original article; instead, we rely on previously-estimated incubation-period distributions by \cite{backer2021omicron} to derive growth-rate-adjusted .
See original article for details of data collection.

@@

In practice, incubation periods (and other epidemiological delays) can be measured in two ways: forward and backward.
The forward incubation periods are measured from a cohort of individuals who were infected at the same time. 
We expect this forward incubation-period distribution $f_I(\tau)$ to remain constant over the course of an epidemic and provide reliable estimates of the distribution across individuals. 
Backward incubation periods are measured from a cohort of individuals who developed symptoms at the same time. 
We expect the backward incubation-period distribution to be sensitive to epidemic dynamics.
The difference arises because forward incubation periods look forward from the reference point towards symptom development, which is an individual-level process, while backward incubation periods look backwards towards an infection event, which requires an interaction with an infectious individual.

@@

For a given growth rate $r$, the corresponding forward incubation-period distributions can be calculated by inverting \eref{backward}:
\begin{equation}
f_I(\tau) = \frac{\exp(r\tau) b_I(\tau) }{\int_0^\infty \exp(rx) b_I(x)\dx}.
\label{eq:forward}
\end{equation}
Since incubation-period data are not provided, we are not able to fit \eref{forward} directly; 
instead we take the backward incubation-period distributions $b_I(x)$ estimated by \cite{backer2021omicron}, which was originally assumed to follow a Weibull distribution, and apply \eref{forward}.
In particular, \cite{backer2021omicron} estimated the scale and shape parameters of the Weibull distribution to be 4.93 (95\% CI: 4.51--5.37) and 1.83 (95\% CI: 1.59--2.08), respectively, for the non-SGTF cases, and 3.60 (95\% CI: 3.23--3.98) and 1.50 ((95\% CI: 1.32--1.70), respectively, for SGTF cases.

We also model the backward incubation-period distribution $b_I(\tau)$ using a Weibull distribution based on the assumptions of \cite{backer2021omicron}.
To account for uncertainties in the original parameter estimates, we rely on a sampling scheme, similar to the one we used for the growth rate analysis (in Section 2.2).
First, we approximate the previously inferred posterior distributions of the shape and scale parameters of the Weibull distribution using a lognormal distribution---we parameterize the lognormal distribution such that (i) its median matches the median of the posterior distributions and (ii) the probability that a random variable following the specified lognormal distribution falls between the lower and upper credible limits matches 95\% \citep{park2020reconciling}.
We draw 1000 samples of the shape and scale parameters (for the backward distribution $b_I(\tau)$) from the specified lognormal distributions and estimate the corresponding forward distribution using \eref{forward}.
We take 95\% equi-tailed quantiles to obtain 95\% confidence intervals.
